Figure two

### 2025-12-05 draftFiguretwo_Ella.Rmd
plots

@@

>>>>

```{r income share}

# 1. Prepare topic-level data (shares within income group, based on unique papers)
topics_data <- papers %>%
  filter(!is.na(Income.group)) %>% 
  distinct(ID, Income.group, category) %>%      # each paper counts once per (Income.group, category)
  count(Income.group, category, name = "n_papers") %>%   # number of papers mentioning category
  rename(Topic = category) %>%
  mutate(Income.group = factor(
    Income.group,
    levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
  ))

# 2. Prepare totals row data (absolute counts of papers per income group)
totals_data <- dataSummary %>%
  filter(!is.na(Income.group)) %>%
  group_by(Income.group) %>% 
  summarize(total = sum(n_papers, na.rm = TRUE), .groups = "drop") %>%
  mutate(Income.group = factor(
    Income.group,
    levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
  ))

topics_data <- topics_data %>%
  left_join(totals_data, by = "Income.group") %>%
  group_by(Income.group) %>%
  mutate(
    share = n_papers / total    # share of papers in that income group mentioning the topic
  ) %>%
  ungroup() %>%
  mutate(Income.group = factor(
    Income.group,
    levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
  ),
  Topic = factor(
    Topic,
    levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
  )
  )

# 3. Define consistent factor levels for the y-axis
# topic_levels <- c("Total papers", unique(topics_data$Topic))

pdf("plots/Figure2_category_paper_by_income_share.pdf", width = 12, height = 10)
ggplot() +
  # --- top totals row ---
  geom_tile(data = totals_data,
            aes(x = Income.group, y = "Total papers", fill = total),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = totals_data,
    aes(x = Income.group, y = "Total papers",
        label = paste0("n=", total),
        color = "black"),
    size = 7
  ) +
  scale_fill_gradient(
    name = "Total papers",
    low = "#F6DBA7", high = "#F7B32B",
    breaks = range(totals_data$total),
    labels = range(totals_data$total)
  ) +
  scale_color_identity() +

  # --- main topic rows ---
  geom_tile(data = topics_data,
            aes(x = Income.group, y = Topic, fill = share),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = topics_data,
    aes(x = Income.group, y = Topic,
        label = scales::percent(share, accuracy = 1),
        # label = paste0(
        #     scales::percent(share, accuracy = 1), 
        #     "\n(", n_papers, ")"
        #   ),
        color = ifelse(share > 0.25, "white", "black")),
    size = 7
  ) +
  scale_fill_gradient(
    name = "Share of papers\nwithin income group ",
    low = "#f7f4f9", high = "#8B475D",
    labels = scales::percent,
    breaks = c(min(topics_data$share), max(topics_data$share))
  ) +
  scale_color_identity() +
  labs(x = "Income group", y = NULL) +
  theme_minimal(base_size = 19) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 22),
    axis.text.y = element_text(size = 22),
    axis.title = element_text(size = 22),
    legend.position = "top"
  )
dev.off()

# 1. Prepare topic-level data (shares within income group, based on unique survey items)
surveys_topics <- surveys %>%
  filter(!is.na(Income.group)) %>%
  mutate(item_id = paste0(iso3, id)) %>%              # unique survey-level ID
  distinct(item_id, Income.group, category) %>%       # each item counts once per (Income.group, category)
  count(Income.group, category, name = "n_items") %>% # number of items mentioning category
  rename(Topic = category) %>%
  mutate(
    Income.group = factor(
      Income.group,
      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
    ),
    Topic = factor(
      Topic,
      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
    )
  )

# 2. Totals per income group (unique items)
surveys_totals <- surveys %>%
  filter(!is.na(Income.group)) %>%
  mutate(item_id = paste0(iso3, id)) %>%
  group_by(Income.group) %>%
  summarise(
    total = n_distinct(item_id),
    .groups = "drop"
  ) %>%
  mutate(
    Income.group = factor(
      Income.group,
      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
    )
  )

surveys_topics <- surveys_topics %>%
  left_join(surveys_totals, by = "Income.group") %>%
  group_by(Income.group) %>%
  mutate(
    share = n_items / total   # share of survey items in that income group mentioning the topic
  ) %>%
  ungroup()

pdf("plots/Figure2_category_survey_by_income_share.pdf", width = 12, height = 10)
ggplot() +
  # --- top totals row ---
  geom_tile(data = surveys_totals,
            aes(x = Income.group, y = "Total items", fill = total),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = surveys_totals,
    aes(x = Income.group, y = "Total items",
        label = paste0("n=", total),
        color = "black"),
    size = 7
  ) +
  scale_fill_gradient(
    name = "Total items",
    low = "#F6DBA7", high = "#F7B32B",
    breaks = range(surveys_totals$total),
    labels = range(surveys_totals$total)
  ) +
  scale_color_identity() +

  # --- main topic rows ---
  geom_tile(data = surveys_topics,
            aes(x = Income.group, y = Topic, fill = share),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = surveys_topics,
    aes(x = Income.group, y = Topic,
        label = scales::percent(share, accuracy = 1),
        # label = paste0(
        #     scales::percent(share, accuracy = 1), 
        #     "\n(", n_papers, ")"
        #   ),
        color = ifelse(share > 0.25, "white", "black")),
    size = 7
  ) +
  scale_fill_gradient(
    name = "Share of items\nwithin income group ",
    low = "#f7f4f9", high = "#541F78",
    labels = scales::percent,
    breaks = c(min(surveys_topics$share), max(surveys_topics$share))
  ) +
  scale_color_identity() +
  labs(x = "Income group", y = NULL) +
  theme_minimal(base_size = 19) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 22),
    axis.text.y = element_text(size = 22),
    axis.title = element_text(size = 22),
    legend.position = "top"
  )
dev.off()

@@

## per exposure level using shares

```{r exposure level shares}

# 1. Prepare topic-level data (shares within exposure group, based on unique papers)
topics_data_poll <- papers %>%
  filter(!is.na(categoryPollution)) %>%
  distinct(ID, categoryPollution, category) %>%          # each paper counts once per (exposure, category)
  count(categoryPollution, category, name = "n_papers") %>%   # number of papers mentioning category
  rename(Topic = category) %>%
  mutate(
    # if you want a specific topic order, same as income plot:
    Topic = factor(
      Topic,
      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
    )
  )

# 2. Prepare totals row data (absolute counts of papers per exposure group)
totals_data_poll <- dataSummary %>%
  filter(!is.na(categoryPollution)) %>%
  group_by(categoryPollution) %>%
  summarize(total = sum(n_papers, na.rm = TRUE), .groups = "drop")

# 3. Add shares within exposure group (based on unique papers)
topics_data_poll <- topics_data_poll %>%
  left_join(totals_data_poll, by = "categoryPollution") %>%
  group_by(categoryPollution) %>%
  mutate(
    share = n_papers / total   # share of papers in that exposure group mentioning the topic
  ) %>%
  ungroup()

# 4. Plot with two independent fill scales
pdf("plots/Figure2_category_paper_by_exposure_share.pdf", width = 12, height = 10)
ggplot() +
  # --- top totals row ---
  geom_tile(
    data = totals_data_poll,
    aes(x = categoryPollution, y = "Total papers", fill = total),
    color = "white", linewidth = 0.8
  ) +
  geom_text(
    data = totals_data_poll,
    aes(
      x = categoryPollution,
      y = "Total papers",
      label = paste0("n=", total),
      color = "black"
    ),
    size = 7
  ) +
  scale_fill_gradient(
    name  = "Total papers",
    low   = "#F6DBA7",
    high  = "#F7B32B",
    breaks = range(totals_data_poll$total),
    labels = range(totals_data_poll$total)
  ) +
  scale_color_identity() +

  # --- main topic rows ---
  geom_tile(
    data = topics_data_poll,
    aes(x = categoryPollution, y = Topic, fill = share),
    color = "white", linewidth = 0.8
  ) +
  geom_text(
    data = topics_data_poll,
    aes(
      x = categoryPollution,
      y = Topic,
      # OPTION 1: percentage only (like your current income plot)
      label = scales::percent(share, accuracy = 1),
      # OPTION 2: uncomment this instead for "percent\n(n_papers)"
      # label = paste0(
      #   scales::percent(share, accuracy = 1),
      #   "\n(", n_papers, ")"
      # ),
      color = ifelse(share > 0.25, "white", "black")
    ),
    size = 7
  ) +
  scale_fill_gradient(
    name   = "Share of papers\nwithin exposure group ",
    low    = "#f7f4f9",
    high   = "#8B475D",
    labels = scales::percent,
    breaks = c(min(topics_data_poll$share), max(topics_data_poll$share))
  ) +
  scale_color_identity() +
  labs(x = "Exposure group", y = NULL) +
  theme_minimal(base_size = 19) +
  theme(
    panel.grid   = element_blank(),
    axis.text.x  = element_text(angle = 30, hjust = 1, size = 22),
    axis.text.y  = element_text(size = 22),
    axis.title   = element_text(size = 22),
    legend.position = "top"
  )
dev.off()

# 1. Prepare topic-level data (shares within exposure group, based on unique survey items)
topics_data_survey_poll <- surveys %>%
  filter(!is.na(categoryPollution)) %>%
  mutate(item_id = paste0(iso3, id)) %>%                      # unique survey item ID
  distinct(item_id, categoryPollution, category) %>%          # each item counts once per (exposure, topic)
  count(categoryPollution, category, name = "n_items") %>%    # number of items mentioning category
  rename(Topic = category) %>%
  mutate(
    # optional: topic order, same as your income-group plot
    Topic = factor(
      Topic,
      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
    )
    # optional: control exposure group order, e.g.:
    # categoryPollution = factor(
    #   categoryPollution,
    #   levels = c("Air pollution", "Water pollution", "Other")  # adjust to your data
    # )
  )

# 2. Prepare totals row data (absolute counts of unique items per exposure group)
totals_data_survey_poll <- surveys %>%
  filter(!is.na(categoryPollution)) %>%
  mutate(item_id = paste0(iso3, id)) %>%
  group_by(categoryPollution) %>%
  summarise(
    total = n_distinct(item_id),
    .groups = "drop"
  )
  # optional: match factor ordering here too
  # %>%
  # mutate(
  #   categoryPollution = factor(
  #     categoryPollution,
  #     levels = c("Air pollution", "Water pollution", "Other")
  #   )
  # )

# 3. Add shares within exposure group (based on unique items)
topics_data_survey_poll <- topics_data_survey_poll %>%
  left_join(totals_data_survey_poll, by = "categoryPollution") %>%
  group_by(categoryPollution) %>%
  mutate(
    share = n_items / total   # share of survey items in that exposure group mentioning the topic
  ) %>%
  ungroup()

# 4. Plot with two independent fill scales
pdf("plots/Figure2_category_item_by_exposure_share.pdf", width = 12, height = 10)
ggplot() +
  # --- top totals row ---
  geom_tile(
    data = totals_data_survey_poll,
    aes(x = categoryPollution, y = "Total items", fill = total),
    color = "white", linewidth = 0.8
  ) +
  geom_text(
    data = totals_data_survey_poll,
    aes(
      x = categoryPollution,
      y = "Total items",
      label = paste0("n=", total),
      color = "black"
    ),
    size = 7
  ) +
  scale_fill_gradient(
    name   = "Total items",
    low    = "#F6DBA7",
    high   = "#F7B32B",
    breaks = range(totals_data_survey_poll$total),
    labels = range(totals_data_survey_poll$total)
  ) +
  scale_color_identity() +

  # --- main topic rows ---
  geom_tile(
    data = topics_data_survey_poll,
    aes(x = categoryPollution, y = Topic, fill = share),
    color = "white", linewidth = 0.8
  ) +
  geom_text(
    data = topics_data_survey_poll,
    aes(
      x = categoryPollution,
      y = Topic,
      # OPTION 1: percentage only (mirroring your income-group code)
      label = scales::percent(share, accuracy = 1),
      # OPTION 2: percentage + underlying n_items
      # label = paste0(
      #   scales::percent(share, accuracy = 1),
      #   "\n(", n_items, ")"
      # ),
      color = ifelse(share > 0.25, "white", "black")
    ),
    size = 7
  ) +
  scale_fill_gradient(
    name   = "Share of items\nwithin exposure group ",
    low    = "#f7f4f9",
    high   = "#541F78",
    labels = scales::percent,
    breaks = c(min(topics_data_survey_poll$share), max(topics_data_survey_poll$share))
  ) +
  scale_color_identity() +
  labs(x = "Exposure group", y = NULL) +
  theme_minimal(base_size = 19) +
  theme(
    panel.grid    = element_blank(),
    axis.text.x   = element_text(angle = 30, hjust = 1, size = 22),
    axis.text.y   = element_text(size = 22),
    axis.title    = element_text(size = 22),
    legend.position = "top"
  )
